\documentclass[12pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage{amssymb}
\title{\vspace{-2.0cm}Determining Non-Negativity of Polynomials using Sum of Squares}
\date{}
\author{Abiyaz Chowdhury, Allen Kim}
\pagenumbering{gobble}
\begin{document}
\maketitle
\vspace{-0.5cm}

The main goal of our project is to implement a sum of squares tester for
arbitrary polynomials. A polynomial $p(x)$ is said to be nonnegative or positive
semidefinite if $p(x) \ge 0$ for all $x \in \mathbb{R}^n$. We also say that a
polynomial $p(x)$ is a sum of squares (sos) if it can be represented as a sum of
square polynomials, meaning that there exists polynomials $q_1(x),\ldots,q_m(x)$
such that $p(x) = \sum_{i=1}^m q_i^2(x)$. It is clear that if a polynomial is
sos, then it is nonnegative. However, the converse is not necessarily true in
general. Determining whether a polynomial is nonnegative is NP-hard even when
the degree of the polynomial is as small as four. However, determining whether a
polynomial has a sos decomposition is doable. In general, this acts as a
tractable substitute for non-negativity.

An important theorem tells us that a multivariate polynomial $p(x)$ in $n$
variables and degree $2d$ is a sum of squares if and only if there exists a
positive semidefinite matrix $Q$ such that $$p(x) = z^T Q z$$ where $z$ is the
vector of monomials of degree up to $d$ $$z = [1,x_1,x_2,\ldots,x_n, x_1x_2,
\ldots, x_n^d]$$

For our project, we aim to determine whether a given polynomial yields a sum of
squares decomposition algorithmically. We outline the steps as follows:
\begin{enumerate}
	\item Convert the given polynomial as a feasibility problem using
	semidefinite program. 
	
	We plan to use SymPy to take an arbitrary polynomial as input. Given the
	coefficients of $p$, we can expand $z^TQz$ and match the coefficients to 
	get linear constraints on $Q$. If we let $q$ be the vectors of $Q$ stacked
	on top of each other vertically, we can represent these linear constraints
	as $Aq = b$. Thus, we get the SDP:
	\begin{center}
	Given a matrix $A$ and vector $b$, find $Q \ge 0$ such that $Aq=b$.
	\end{center}

	\item Solve the semidefinite program using interior point methods.

	We hope to implement some interior point method using log barrier functions
	to solve the SDP.

	\item Convert the solution to the SDP back into a sum of squares (if one
	exists).

	Once we have a $Q$, since $Q$ is a PSD matrix, we can compute its Cholesky
	decomposition to get $Q = V^TV$ and then, output the sum of squares
	directly: $$p(x) = \sum_i (Vz)_i^2

\end{enumerate}

\end{document}

